---
title: "Review of: Unprecedented disease-related coral mortality in Southeastern Florida"
author: "Rachel N. Silverstein, Ross Cunning, Andrew C. Baker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Introduction
This document attempts to replicate results presented in Precht et al. (2016) in *Scientific Reports* [(doi:10.1038/srep31374)](http://doi.org/10.1038/srep31374) using the dataset provided to us by the authors upon request on March 8, 2018. After importing the data, we proceed line by line through the 'Tagged-coral colonies' section of the results, with block quotes taken directly from the paper, followed by our attempt to reproduce each result. For full transparency and reproducibility, this document contains the R code we used in each attempt to reproduce reported results, along with output in the form of figures and tables. Finally, we describe and explain the outcome of our attempt to reproduce the authors' reported findings.

In short, we are unable to reproduce many findings reported in the tagged coral Results section of the manuscript using the dataset provided by the authors. Some discrepancies in the data are more dramatic than others. Upon inquiry with the authors, we now understand a major factor in our inability to reproduce the disease prevalence data was that the data was further changed by the authors in a manner not described in the Methods -- and which differs from the dataset initially provided to us. An extra spreadsheet indicating additional corals that authors had deemed as "white plague mortality" post hoc was subsequently provided to us on April 4, 2018. Communication with authors and information provided in the additional spreadsheet indicates that changes were made to the data based on a set of assumptions that are not reflected in the dataset initially provided, and therefore cannot be reproduced. The data for disease prevalence reported in the Results, for example, were modified to include (1) corals that died possibly weeks to months prior than the date in question but that previously had white plague disease or was assumed to have white plague disease, and (2) "assumed" disease mortality. The authors, through a set of criteria not described in the Methods, assume that corals died from white plague disease without ever being observed with disease symptoms. Even with this additional spreadsheet, however, we still cannot reproduce all of Precht et al's reported findings. Data issues are described in detail below.

##### Background on data and sampling design

Precht et al. 2016 includes coral condition observation data from both timed swims and tagged corals. The tagged coral data is the focus of our analysis as it was collected as a part of a Florida Department of Environmental Protection (FDEP) permit-mandated compliance monitoring connected to the Port of Miami dredging project, which took place from October 2013 through September 2015. Three of four authors are also employees of Dial Cordy and Associates, subcontractors of the dredging company Great Lakes Dredging and Dock, who conducted this compliance monitoring. 

Compliance monitoring was triggered whenever dredging activity was within 750m of the monitoring locations.  There were a total of 26 tagged coral monitoring locations located north and south of the shipping channel and sites near the channel (wihtin 100 ft) and far from the channel at paired "control" sites (ranging from several hundred meters to the south to thousands of meters from the channel to the north). Reef areas being monitored included hardbottom areas (closest inshore), the middle reef, and the outer reef areas (together these reefs comprise the Florida Reef Tract).

Precht et al. only reports data from 2 paired control sites on the northern and 2 paired control sites on the southern middle reef, which represents only 4 of 26 permanement montioring stations in which the authors had tagged corals. They do not include their data from the outer reef or "hardbottom" sites, or any sites near the shipping channel where dredging was occurring.

Authors do not mention additional tagged coral data or the nearby dredging project. Observations reported in Precht et al. span the entire course of the dredging project from October 2013 through July 2015, a few months after dredging concluded. The control locations from which they report data to the south of the channel were impacted by dredging sediment (P. Swart, 28 June 2016, "Report on the Mineralogy and the Stable Carbon and Oxygen Isotopic Composition of Samples Supplied by NOAA"). 
 
A full dataset from all 26 monitoring locations including reef & hardbottom sites (northern and southern, middle and outer reefs, channel-side and control locations) was also submitted to the FDEP for permit compliance monitoring. We requested the full monitoring dataset in a public records request from FDEP (“FDEP-provided data”). We find many discrepancies between the data that the authors submitted to FDEP and in the dataset that they provided to us for to support their findings. Analysis of these discrepancies is available upon request.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load packages
library(tidyverse)
library(stringr)
library(lubridate)
library(readxl)
library(janitor)

# Function for accessing data for a specific coral from full dataset
getCoral <- function(d, s, t, c) {
  with(d, d[Site==s & Transect==t & Coral.ID==c, ])
}

# Function for accessing corals with specific condition codes
hasCond <- function(cond, Condition.Codes) {
  unlist(lapply(Condition.Codes, FUN=function(x) any(grepl(cond, x))))
}
```

## 1. Import and organize data
```{r load_scirep_data}
scirep <- read_xlsx("POM_R2_ControlData_Baseline_Post_construction.xlsx")

# Tidy data
names(scirep) <- c("Date", "Week", "Visit", "Transect", "Category", "Species",
                     "Coral.ID", "QAQC", "Condition.Code")
scirep <- scirep %>%
  mutate(Date = as.Date(Date),
         Week = as.factor(Week)) %>%
  separate(Transect, into = c("Site", "Site.2", "Transect")) %>%
  select(Date, Week, Site, Transect, Coral.ID, Species, QAQC, Condition.Code) %>%
  mutate(coral = interaction(Site, Transect, Coral.ID)) %>%
  filter(Species != "Tag not found") %>%
  droplevels()

# Identify corals recorded as WP prior to being recorded as DEAD
wpdead <- scirep %>%
  filter(Condition.Code %in% c("WP", "DEAD")) %>%
  group_by(coral) %>%
  summarise(firstWP = as.character(min(Date[Condition.Code=="WP"], na.rm=T)),
            firstDEAD = as.character(min(Date[Condition.Code=="DEAD"], na.rm=T))) %>%
  drop_na() %>%
  mutate(WPbeforeDEAD = ifelse(as.Date(firstDEAD) > as.Date(firstWP), TRUE, FALSE))

# For these corals recorded as WP prior to DEAD, change DEAD to WPDEAD
wpdead2 <- scirep %>%
  filter(coral %in% wpdead$coral,
         Condition.Code=="DEAD") %>%
  mutate(Condition.Code="WPDEAD")

scirep <- bind_rows(scirep, wpdead2)

# Merge condition codes
scirep <- with(scirep, {
  aggregate(data.frame(Condition.Codes=Condition.Code), 
            by=list(Date=Date, Week=Week, Site=Site, Transect=Transect, 
                    Coral.ID=Coral.ID, Species=Species, coral=coral), 
            FUN=function(x) c(as.character(x)))
})
```

#### 1.1 Identify species ID errors for individual tagged corals

Here we check to see whether the species IDs associated with each tagged coral are the same across all observations.

```{r check_conflicts}
# How many corals have multiple species assignments? (e.g., indicative of an error in data entry)
sppcalls <- aggregate(data.frame(Species=scirep$Species), list(Site.Transect.CoralID=scirep$coral), 
                      FUN=function(x) unique(c(as.character(x))))

# Save conflicting species assignments to file
capture.output(sppcalls, file="precht_spp_conflicts.txt")

# Count number of corals with multiple conflicting species assignments
conflicts <- table(sapply(sppcalls$Species, length) > 1, dnn = list("conflicting"))

knitr::kable(as.data.frame(conflicts, responseName = "n"))
```

Out of 115 corals, 23 have multiple species IDs associated with them at different points in time, indicative of some (unidentifiable) error in data entry. A full list of these conflicting species assignments is provided in the attached file: "precht_spp_conflicts.txt".

#### 1.2 Change conflicting species assignments to reflect most frequently assigned species

We attempt to fix these conflicting species assignments in the present analysis by changing species assignments for each coral (defined as unique site-transect-coralID) to the most frequently assigned species.

```{r fix_errors_scirep}
# There are several instances where a given Site-Transect-Coral.ID has multiple spp. assignments...
# Find which name was used most frequently for each Site-Transect-Coral.ID
which.spp <- with(scirep, {
  aggregate(data.frame(Species=Species), by=list(Site=Site, Transect=Transect, Coral.ID=Coral.ID), 
            FUN=function(x) {
              names(table(c(as.character(x))))[which.max(table(c(as.character(x))))]
            })
})

# Reassign all values of Species for each coral to the most frequently used species
for (i in 1:nrow(which.spp)) {
  scirep[scirep$Site==which.spp[i, "Site"] & 
          scirep$Transect==which.spp[i, "Transect"] &
          scirep$Coral.ID==which.spp[i, "Coral.ID"],
          "Species"] <- which.spp[i, "Species"]
}
```

## 2. Dataset summary

> A total of 115 coral colonies, representing 13 coral species, were tagged for repeated monitoring at four sites on the inner reef tract in Miami-Dade County (Fig. 1). Coral colonies were monitored at least 40 times between October 19, 2013 and July 13, 2015.

```{r n_tagged} 
# Count number of corals of each species at each site
ncorals <- scirep %>%
  distinct(Site, Species, coral) %>%
  count(Site, Species) %>%
  spread(Species, n) %>%
  adorn_totals()
  
knitr::kable(ncorals, caption = "Number of tagged corals")
```

Number of times monitored:  

```{r n_times_monitored, fig.width=4, fig.height=2}
# Count number of times each coral was observed
n_times <- scirep %>%
  group_by(coral) %>%
  summarise(n = nlevels(factor(Date)))

# Plot histogram of times observed
ggplot(n_times, aes(x = n)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = seq(30, 50, 1)) +
  labs(x = "Number of times observed", y = "Number of corals")
```

Within the dataset provided, we do find that 115 corals of 13 species were monitored. However, we found that individual colonies were monitored between `r min(n_times$n)` and `r max(n_times$n)` times each. This is inconsistent with the statement that colonies were observed at least 40 times.

------


## 3. Peak of coral bleaching

> The peak of coral bleaching was recorded in September 2014. Specifically, the highest recorded bleaching prevalence occurred on September 12, 2014, when 84% (21 of 25 corals surveyed that day) showed signs of coral bleaching (Fig. 2).  

To reproduce this result, we will look at the three different condition codes: BL, PB, and P. In the definitions of the condition codes prescribed by FDEP protocols, P ('paling') is defined as live tissue with some loss of color, PB ('partial bleaching') is defined as “patches of fully bleached or white tissue”, and BL ('bleaching') is defined as “live tissue with complete loss of color across the entire colony.” 

```{r peak_bleaching}
# Quantify bleaching (condition codes BL, PB, and/or P) on 9/12/14
scirep %>%
  filter(Date=="2014-09-12") %>%
  summarise(BL = sum(hasCond("^BL$", Condition.Codes)),
            `BL or PB` = sum(hasCond("^PB$|^BL$", Condition.Codes)),
            `BL or PB or P` = sum(hasCond("^PB$|^BL$|^P$", Condition.Codes)),
            `(Total)` = n()) %>%
  gather(key = "Condtion Code", value = "Number of corals") %>%
  mutate(Proportion = `Number of corals` / 25) %>%
  knitr::kable()
```

We are able to reproduce the 84% metric reported in the paper, but only if all corals recorded with the condition codes for “paling” (P), “partial bleaching” (PB), or “bleaching” (BL) are combined together. The proportion of true "bleaching" (BL) in the condition code data was 4 corals out of 25 (16%). Partial bleaching (PB) condition code accounts for an additional 3 corals out of 25. The majority of the 84% "bleaching", reported here, therefore, is just due to paling. Combining these condition codes to describe “bleaching” is misleading and serves to inflate the apparent bleaching prevalence.

Here, we tally and plot the prevalence of all three of these condition codes over time for the entire dataset:

```{r peak_bleaching2_scirep, fig.width=10, fig.height=3}
bleaching <- scirep %>%
  group_by(Date) %>%
  summarise(n=n(),
            freq.BL=sum(hasCond("^BL$", Condition.Codes))/n,
            freq.BLorPB=sum(hasCond("^BL$|^PB$", Condition.Codes))/n,
            freq.BLorPBorP=sum(hasCond("^BL$|^PB$|^P$", Condition.Codes))/n)

# Plot frequency of bleaching codes recorded over time            
ggplot(data=bleaching, mapping=aes(x=Date)) + ylab("Frequency") +
  geom_bar(aes(y=freq.BLorPBorP), stat="identity", fill="green") +  # Green is BL, PB, or P
  geom_bar(aes(y=freq.BLorPB), stat="identity", fill="blue") +  # Blue is BL or PB
  geom_bar(aes(y=freq.BL), stat="identity", fill="red") + # Red is BL only
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

We find that the maximum prevalence of bleaching indicated by the BL condition code (red) was September 15th (20% of observed colonies), and the maximum prevalence of bleaching and partial bleaching (BL or PB condition codes -- blue) was on September 27th (43% of observed colonies).

------

## 4. First signs of white plague

> The first sign of a white-plague disease outbreak (prevalence > 5%) was noticed at the southern monitoring sites, near Virginia Key, on September 26, 2014.

```{r wp_frequency}
# Calculate frequency of WP at each Site on each Date
wp_scirep <- scirep %>%
  group_by(Site, Date) %>%
  summarise(n = n(),
            WP = sum(hasCond("^WP$", Condition.Codes)),
            freq.WP = WP / n)

# What was the frequency of WP at the southern monitoring sites on September 26, 2014?
wp_scirep %>% filter(Date=="2014-09-26") %>% knitr::kable()
```

Using the data provided by the authors, we find that the prevalence of WP on Sept. 26 2014 was not greater than 5% at southern monitoring sites.
 
Disease prevalence was only 4% at site R2SC2 (N=1 out of 25 corals) and was 0% at R2SC1 (R2SC2 and R2SC1 are the southern monitoring sites referenced in the Results statement above). Neither reached the “5% prevalence” threshold on that date.

In response to our inquiry regarding this discrepancy, the authors responded that "the original statement should have read '< 5%' not '> 5%.'" This "typo" fundamentally changes the conclusions of the paper and the contagion-based model presented in the paper.

```{r first_wp_over5}
# Find earliest Date with WP frequency greater than 5% at each Site
wp_scirep %>%
  group_by(Site) %>%
  summarise(`First WP > 5%` = min(Date[freq.WP > 0.05]),
                          n = n[Date == `First WP > 5%`],
                         WP = WP[Date == `First WP > 5%`],
                    freq.WP = freq.WP[Date == `First WP > 5%`]) %>%
  knitr::kable()
```

In the 4 monitoring sites included in Precht et al., we find that the first time >5% white plague prevalence was documented was on October 17, 2014 at R2SC2, when proportion was 8% (N=2 out of 25 colonies).
 
However, considering all 25 permanent monitoring locations (data submitted to FDEP as part of compliance monitoring), we find that the first prevalence >5% WP at a single site was on June 27th 2014 at R3NC1 (the northern control site from the outer reef; outer reef sites were all omitted from the analyses presented in this paper). This puts the 'start of the outbreak' (indicated by >5% prevalence) months earlier and before bleaching had occurred, and north of the channel where dredging was occuring. We find that the first report of any white plague disease at all (even <5%) occurred on October 23, 2013 at site R2N1 (northern middle reef channel-side site), which was also omitted from the paper.
 
For Authors' response to our inquiry about this issue, see “Addl Comments on Response to Rachel Silverstein- Part 2, page 4".

------

## 5. White plague prevalence and dynamics

> ...by November 18, 15% of the tagged colonies showed outward signs of white-plague disease (Fig. 2).

```{r wp_2014-11-18}
# How many colonies had WP on November 18th, 2015?
scirep %>%
  filter(Date=="2014-11-18") %>%
  summarise(Date=unique(Date),
            n=n(),
            WP=sum(hasCond("^WP$", Condition.Codes)),
            freq.WP=WP/n) %>%
  knitr::kable()
```

We find that only 6 of 55 colonies (10.9%) observed on November 18th were recorded with the WP condition code.

```{r wp_or_dead_2014-11-18}
# How many colonies had WP or were DEAD on November 18th, 2015?
scirep %>%
  filter(Date=="2014-11-18") %>%
  summarise(Date=unique(Date),
            n=n(),
            WPorDEAD=sum(hasCond("^WP$|^WPDEAD$", Condition.Codes)),
            freq.WPorDEAD=WPorDEAD/n) %>%
  knitr::kable()
```

We are unable to reproduce the 15% number reported in the study, even if corals that had been recorded with white plague in the past and then died are counted (Prevalence of WP or WPdead= 12.7%, 7 out of 55 colonies observed). Based on communication with the authors (see "Response to SciRep questions from RS.pdf"), it appears that the remaining 2.3% of corals were assumed to have died from WP without ever having observed the coral with white plague disease. Authors stated that they ascribed additional corals that were found dead as having had white plague using a set of criteria that is not described in the paper[^1] and is not reflected in their dataset. The reported value therefore does not describe the colonies that “showed outward signs of white plague” on this date.

------

> White-plague disease did not impact the northern monitoring sites (14 km north) until June 2015.

```{r wp_north}
# When was WP first observed at ≥5% prevalence at the northern sites?
wp_scirep %>%
  filter(Site %in% c("R2NC1", "R2NC2")) %>%
  group_by(Site) %>%
  summarise(`First WP > 5%` = min(Date[freq.WP > 0.05]),
                          n = n[Date == `First WP > 5%`],
                         WP = WP[Date == `First WP > 5%`],
                    freq.WP = freq.WP[Date == `First WP > 5%`]) %>%
  knitr::kable()
```

We also find that WP was first observed ≥5% at the northern sites in June 2015 within the provided dataset. However, data provided to FDEP for compliance monitoring from the same 4 sites shows that a northern control site (R2NC2) was directly observed to have a 10% WP prevalence on May 19, 2015. However, Precht et al. excludes observations from this date in their analysis. For the authors' response to our inquiry about this missing date, see "Response to Rachel Silverstein- Part 2 (WFP & BEG 4.20.18).pdf"
 
The exclusion of over 75% of the monitoring sites where the authors were actively collected tagged coral data during this time, as well as excluding certain datapoints in time, fundamentally alters the apparent appearance and spread of the disease.

------

> During the week of June 17 [2015], 17% of the tagged coral colonies at the northern monitoring sites were infected with white-plague disease.

```{r wp_2015-06-17}
# How many corals at northern sites had WP on June 17, 2015?
scirep %>%
  filter(Site %in% c("R2NC1", "R2NC2"), Date=="2015-06-17") %>%
  summarise(Date=unique(Date),
            n=n(),
            WP=sum(hasCond("^WP$", Condition.Codes)),
            freq.WP=WP/n) %>%
  knitr::kable()
```

We find that out of 60 corals at the northern monitoring sites, 6 (10%) were directly observed with active white plague on June 17th, 2015. In order reproduce the 17% prevalence reported here, authors must have included corals that were found dead and were assumed to have died from white plague disease but without direct observation or ever being recorded as having white plague [see footnote 1].
 
Therefore, this result is not reproducible and is misleading -- almost half of these corals were not actually "infected with white-plague disease" on this date.

[^1]: The authors described their method of attributing white plague as the cause of mortality without directly observing the disease on colonies as follows: “All corals that were included in the accounting of the effects of WP disease that were not directly observed with active disease were included if: 1) the coral was a WP susceptible species, 2) total colony mortality occurred during the WP disease epidemic timeframe, 3) WP was active and killing susceptible corals at the survey site, and 4) no other obvious signs or causes of mortality (i.e. competition, white band disease, black band, etc.) were noted.” (via email attachment “Response to Sci Rep Questions from RS” on April 4, 2018)

------

> In total, white-plague disease was observed on seven of the 13 coral species identified at the four permanent monitoring sites (Table 1). 

```{r wp_spp}
# How many species were observed to ever have WP?
scirep %>%
  group_by(Species) %>%
  summarise(`WP observed`=any(hasCond("^WP$", Condition.Codes))) %>%
  knitr::kable()
```

Based on the Precht et al. dataset, only 6 species are shown to have WP at the four permanent monitoring sites. The data do not contain any reports of WP for *Colpophyllia natans*, although this species is reported as having the disease in the manuscript (Precht et al. Table 1). Photographs provided by FDEP show that the single tagged *C. natans* colony was indeed diseased on Dec. 12, 2014, but data from this date are not present in the authors' provided dataset and therefore the analysis is not reproducible [^2].

[^2]: Data appears in the dataset from December 12, 2014 but only from R2SC1 location (southern control site). Photographs submitted to FDEP show that corals at R2SC2 were observed that date but were not recorded in the dataset.

------

> The highest recorded prevalence of white-plague disease occurred on July 7, 2015, when 40% (22 of 55 corals surveyed that day) showed signs of white-plague disease infection (Fig. 2). 

```{r wp_2015-07-07, fig.width=10, fig.height=3}
# What is prevalence of WP on July 7th, 2015?
scirep %>%
  filter(Date=="2015-07-07") %>%
  summarise(Date=unique(Date), n=n(), 
            WP=sum(hasCond("^WP$", Condition.Codes)), 
            freq.WP=WP/n) %>%
  knitr::kable()

scirep %>%
  filter(Date=="2015-07-07") %>%
  summarise(Date=unique(Date), n=n(), 
            WPorWPDEAD=sum(hasCond("^WP$|^WPDEAD$", Condition.Codes)), 
            freq.WPorWPDEAD=WPorWPDEAD/n) %>%
  knitr::kable()
```

We find that on July 7, 2015, only 3 out of 55 corals observed (5.5%) were recorded as having WP. If we include corals that were recorded as having WP *and* corals that were previously recorded as having WP and are now recorded as dead, the total is 15 out of 55 corals, or 27.3%. Therefore, we are unable to reproduce the result reported in the paper with the provided data. Based on responses to inquiry, authors provided an additional dataset showing an additional 7 dead corals that they believe died from white plague infection, though they are never directly observed with the disease. To reach the reported 40%, these corals, too, are counted as showing "signs of white-plague disease infection” on July 7, 2015. Describing  dead corals as "showing signs of white plague infection" using this value to describe disease prevalence is not accurate and overinflates the apparent impact of the disease.
 
When was the highest recorded prevalence of active white plague?

```{r scirep_highest_wp_prevalence, fig.width=10, fig.height=3}
# Graph white plague prevalence by Date (for all corals surveyed - may include multiple sites)
scirep_wp_by_date <- scirep %>%
  group_by(Date) %>%
  summarise(n=n(), WP=sum(hasCond("^WP$", Condition.Codes)), freq.WP=WP/n)

ggplot(data=scirep_wp_by_date, aes(x=Date, y=freq.WP)) + ylim(0,0.4) + 
  geom_bar(stat="identity") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Of only the tagged corals directly observed to have white plague disease on a given date, we find the highest active disease prevalence only ever reaches 15% at a single site (9 out of 60 colonies observed on July 5, 2015).

We can also show the cumulative impact of white plague disease by plotting the number of corals with active white plague combined with corals that had previously been observed with white plague disease and are now dead:

```{r, fig.width = 10, fig.height = 3}
# Graph prevalence of WP or WPDEAD
scirep_wporwpdead_by_date <- scirep %>%
  group_by(Date) %>%
  summarise(n=n(), WPorWPDEAD=sum(hasCond("^WP$|^WPDEAD$", Condition.Codes)), 
            freq.WPorWPDEAD=WPorWPDEAD/n)

ggplot(data=scirep_wporwpdead_by_date, aes(x=Date, y=freq.WPorWPDEAD)) + ylim(0,0.4) +
  geom_bar(stat="identity") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

If we count corals that have active white plauge disease (WP) *and* corals that previously had WP disease and are now dead, we find that the highest prevalence on any date is 15 out of 55 (27.3%) on July 7th, 2015.

------

> Of the white-plague impacted species, the overall disease prevalence was 51% (35 of 69 corals surveyed).

```{r wp_freq_sus_spp}
# Susceptible species:
sus <- c("CNAT", "DSTO", "DLAB", "EFAS", "MMEA", "MCAV", "ODIF", "DCLI", "DSTR", "SBOU")

scirep_wp.corals <- scirep %>%
  filter(Species %in% sus) %>%
  group_by(Species, coral) %>%
  summarise(WP=any(hasCond("^WP$", Condition.Codes))) %>%
  select(WP) %>% 
  table() %>%
  addmargins() %>%
  knitr::kable()
```

We find that the dataset provided shows that only 28 out of 69 colonies (40.6%) were recorded as having WP at any point in time. Therefore, the reported disease prevalence of 51% must also include corals that were found dead and were assumed to have died from white plague.

------

## 6. Relationship between bleaching and disease

------

> The majority (81%) of white-plague susceptible corals bleached, prior to becoming infected with the disease (Table 1). *Meandrina meandrites*, *Dichocoenia stokesi*, *Colpophyllia natans*, and *Pseudodiploria strigosa* were the most heavily impacted coral species, with 100% of tagged colonies being afflicted with white-plague disease. All colonies that showed signs of disease died, regardless of species. 

```{r scirep_bleaching_disease}
# Get list of WP-affected corals
scirep_wp.corals <- scirep %>%
  filter(Species %in% sus) %>%
  group_by(coral) %>%
  summarise(WP=any(hasCond("^WP$", Condition.Codes))) %>%
  filter(WP==TRUE)

# How many WP-affected corals bleached prior to becoming infected?
scirep_bleach_dis <- scirep %>%
  filter(coral %in% scirep_wp.corals$coral) %>%
  group_by(coral) %>%
  summarise(`P date`=min(Date[hasCond("^P$", Condition.Codes)]),
            `PB date`=min(Date[hasCond("^PB$", Condition.Codes)]),
            `BL date`=min(Date[hasCond("^BL$", Condition.Codes)]),
            `WP date`=min(Date[hasCond("^WP$", Condition.Codes)])) %>%
  arrange(`BL date`, `PB date`, `P date`, `WP date`)

# Calculate number of corals with either BL, PB, or P prior to WP
scirep_bleach_dis %>%
  summarise(`n with WP`=n(),
            `BL first`=sum(`BL date` < `WP date`),
            `PB first`=sum(`PB date` < `WP date`),
            `P first`=sum(`P date` < `WP date`)) %>%
  knitr::kable()
```

We cannot reproduce this finding. We find that only 3 out of 28 corals (10.7%) that had WP were recorded as bleached (BL) prior to becoming infected. 20 out of 28 (71%) were recorded as partially bleached (PB) prior to infection, while 28 out of 28 (100%) were recorded as pale (P) prior to infection. Therefore, all three conditions codes could not have been combined to report this finding, as they were in earlier reports of bleachign prevalence and we cannot replicate 81% prior bleaching for WP-infected corals.

However, to statistically test whether there is a relationship between bleaching and disease, the number of corals that bleached and did NOT get disease must also be considered.

```{r bleaching_nodisease}
blwp <- scirep %>%
  filter(Species %in% sus) %>%
  group_by(coral) %>%
  summarise(firstPBorBL = min(Date[hasCond("^PB$|^BL$", Condition.Codes)]),
                firstWP = min(Date[hasCond("^WP$", Condition.Codes)])) %>%
  mutate_if(is.Date, function(x) (x - as.Date("2013-10-19")) / ddays(1)) %>%  # convert to days since start
  mutate(firstPBorBL = if_else(is.infinite(firstPBorBL), 999, firstPBorBL),
             firstWP = if_else(is.infinite(firstWP), 999, firstWP)) 

res <- with(blwp, table(WP = firstWP < 999, BLorPB = firstPBorBL < firstWP))
addmargins(res)

fisher.test(res)
```

Because we do not consider "paling" to be descriptive of bleaching, we considered BL and PB only. We find that 50 of the 69 disease-susceptible corals 'bleached' (BL or PB), and of these, 20 got WP disease (40%), but 30 did not (60%). We find a similar incidence of disease among corals that did not bleach, with 8 out of 19 (42%) having the disease, and 11 (58%) not having disease. With Fisher's exact test, we find no relationship between bleaching (BL or PB) and disease (WP) (p-value=1).

------

## 7. Relationship between disease and death
 
> All colonies that showed signs of disease died, regardless of species.

```{r}
scirep %>%
  group_by(coral) %>%
  summarise(WP = any(hasCond("^WP$", Condition.Codes)),
            DEAD = any(hasCond("^DEAD$", Condition.Codes))) %>%
  count(WP, DEAD) %>%
  knitr::kable()
```

We find that 15 corals that were recorded as WP were still alive on 7/15/15. Therefore, the data do not support the claim of 100% mortality following disease infection.
 
In responses to our questions, the authors also confirm that at least 1 white-plague infected coral at the 4 sites analyzed was still alive during monitoring`, saying “… a tagged coral at R2NC1 was noted with WPD on 8/13/2014 but the colony remained alive through active construction monitoring..." (See Response to Rachel Silverstein - Part 2 (WFP & BEG 4.20.18).pdf"
