---
title: "partial_mortality.Rmd"
author: "Ross Cunning"
date: "5/23/2018"
output: html_document
---

```{r setup, include=FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load packages
library(tidyverse)
library(lme4)
library(lsmeans)

# Define function for accessing corals with specific condition codes
hasCond <- function(cond, Condition.Codes) {
  unlist(lapply(Condition.Codes, FUN=function(x) any(grepl(cond, x))))
}
```

# Import data
```{r}
# Import tidy tagged coral data (does not include visual estimates of partial mortality extent)
load("../data/tagged_corals.RData")

# Import 2016 visual estimate of extent of partial mortality data
ia <- read_csv("../data/DEP_responsive_records_May_2017/permanent_sites_impact_assessment_corals_2016_CSI_QAQC.csv", skip=1) %>%
  select("Date", "Site", "Transect", "Tagged coral (Y/N)", "Coral ID", "Species", "Visual Estimate % Mortality") %>%
  filter(`Tagged coral (Y/N)` == "Y") %>%
  rename("pct.mort" = "Visual Estimate % Mortality") %>%
  mutate(coral = interaction(Site, Transect, `Coral ID`),
         Date = as.Date(Date, format = "%m/%d/%y")) %>%
  select("Date", "coral", "Species", "pct.mort") %>%
  filter(grepl("^R", coral)) %>%
  mutate(pct.mort = as.numeric(pct.mort)) %>%
  distinct()  # remove duplicated rows


# Merge percent mortality with rest of data
pmor <- left_join(ia, alldata, by=c("Date", "coral", "Species")) %>%
  select(Date, reef, dir, Site, Channel, Transect, Coral.ID, Species, pct.mort, coral)
pmor <- pmor[!is.na(pmor$Channel),]  # A couple of spp IDs didn't match (data entry error?), so other columns did not merge. omit these rows.
```

# Subset non-disease-susceptible corals
```{r}
# WP disease-susceptible species
sus.spp <- c("CNAT", "DSTO", "DLAB", "EFAS", "MMEA", "MCAV", "ODIF", "DCLI", "DSTR", "SBOU")

pmor <- filter(pmor, !Species %in% sus.spp)  # exclude susceptible corals
```

# Analyze 2016 visual estimates of partial mortality
### Extent of partial mortality for each coral by site
```{r, fig.width = 10, fig.height = 3}
# Analyze loss of tissue
ggplot(pmor, aes(x = Site, y = pct.mort)) + geom_violin() + geom_jitter()
```

### Test differences in partial mortality between channelside and control sites
```{r}
# Fit model
pmor.mod1 <- pmor %>%
  glm(pct.mort/100 ~ Channel * reef * dir, data = ., family = "quasibinomial")

# Get lsmeans
pmor.lsm1 <- lsmeans(pmor.mod1, specs = c("Channel", "reef", "dir"), type = "response")

# Test for differences between channelside and control for each reef/direction
rbind(pairs(pmor.lsm1, by = c("reef", "dir")), adjust = "none")

# Plot lsmeans
plot(pmor.lsm1, main = "% partial mortality (non-disease-susceptible corals)")
```

Non-disease-susceptible corals on the northern middle reef (R2 N) showed increased percent partial mortality at channelside sites (48.3%) relative to control sites (17.5%). There was also a significant difference for the southern middle reef, but control sites had higher partial mortality (35.8%) than the channelside sites (8.5%).

### Test individual species at north middle reef
```{r}
pmor.mod2 <- pmor %>%
  filter(reef == "R2", dir == "N") %>%
  glm(pct.mort/100 ~ Channel * Species, data = ., family = "quasibinomial")

# Get lsmeans
pmor.lsm2 <- lsmeans(pmor.mod2, specs = c("Channel", "Species"), type = "response")

# Pairwise comparisons (channelside vs. control) for each species
rbind(pairs(pmor.lsm2, by = c("Species")), adjust = "none")

# Plot lsmeans
plot(pmor.lsm2)
```

*Siderastrea siderea* has significantly higher partial mortality at channelside sites (53.6%) compared to control sites (8.3%) on the northern middle reef. For *Stephanocoenia intersepta*, partial mortality is also higher at channelside relative to control sites (41.0% vs. 7.5%), but this difference was not significant. For *P. astreoides*, there were no tagged corals at the channelside sites at the northern middle reef.

# Analyze PM condition code over time