---
title: "Import and tidy tagged coral data"
author: "Ross Cunning, Rachel Silverstein"
date: "3/07/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r setup2}
# Load packages
library(tidyverse)
library(stringr)
library(lubridate)
library(readxl)

# Function for accessing data for a specific coral from full dataset
getCoral <- function(d, s, t, c) {
  with(d, d[Site==s & Transect==t & Coral.ID==c, ])
}

# Function for accessing corals with specific condition codes
hasCond <- function(cond, Condition.Codes) {
  unlist(lapply(Condition.Codes, FUN=function(x) any(grepl(cond, x))))
}
```

# Import data

### Baseline

##### Hardbottom
```{r baseline_hb}
# Import data from file: Hardbottom_Baseline_POM_050714.xlsx
bl_hb <- readxl::read_xlsx(path = "data/to_analyze/Hardbottom_Baseline_POM_050714.xlsx",
                           sheet = "Raw Data")
# Tidy data
bl_hb <- bl_hb %>%
  clean_names() %>%
  filter(tagged_coral_y_n == "Y") %>%         # Tagged corals only
  rename(site = site_e_g_hbnc1_cp,
         transect = transect_e_g_hbna_1,
         species = subcategory) %>%
  mutate(date = as_date(date)) %>%
  select(date, site, transect, coral_id, species, condition_qa_qc) %>%
  rename(condition = condition_qa_qc)

# Check and make sure no missing dates
#bl_hb %>% filter(is.na(date))

# Separate condition codes into multiple rows
bl_hb <- bl_hb %>%
  separate_rows(condition, sep = "/")
```

##### Middle reef
```{r baseline_r2}
# Import data from file: Reef2_Baseline_POM_2013_050714.xlsx
bl_r2 <- readxl::read_xlsx(path = "data/raw/Reef2_Baseline_POM_2013_050714.xlsx",
                           sheet = "Raw Data", guess_max = 3000)

# Tidy data
bl_r2 <- bl_r2 %>%
  clean_names() %>%
  filter(tagged_coral_y_n == "Y") %>%
  rename(site = site_e_g_r2n1_rr,
         transect = transect_e_g_r2n1_rr_1,
         species = subcategory) %>%
  mutate(date = as_date(date)) %>%
  select(date, site, transect, coral_id, species, condition_qa_qc) %>%
  rename(condition = condition_qa_qc)

# Check and make sure no missing dates
#bl_r2 %>% filter(is.na(date))
# Correct date errors based on manual inspection of spreadsheet
bl_r2 <- bl_r2 %>% 
  mutate(date = case_when(
    transect=="R2N1-RR-1" & coral_id==10 & is.na(date) ~ as_date("2013-11-02"),
    transect=="R2NC3-LR-2" & coral_id==9 & is.na(date) ~ as_date("2013-11-19"),
    TRUE ~ as_date(date)))

# Separate condition codes into multiple rows
bl_r2 <- bl_r2 %>%
  separate_rows(condition, sep = "/")
```

##### Outer reef
```{r baseline_r3}
# Import data from file: Reef2_Baseline_POM_2013_050714.xlsx
bl_r3 <- readxl::read_xlsx(path = "data/raw/Reef3_Baseline_POM_050714 (REVISED DOCUMENT).xlsx",
                           sheet = "Raw Data", guess_max = 3000)

# Tidy data
bl_r3 <- bl_r3 %>%
  clean_names() %>%
  filter(tagged_coral_y_n == "Y") %>%
  rename(site = site_e_g_hbnc1_cp,
         transect = transect_e_g_hbna_1,
         species = subcategory) %>%
  mutate(date = as_date(date)) %>%
  select(date, site, transect, coral_id, species, qa_qc_condition_code) %>%
  rename(condition = qa_qc_condition_code)

# Check and make sure no missing dates
#bl_r3 %>% filter(is.na(date))

# Separate condition codes into multiple rows
bl_r3 <- bl_r3 %>%
  separate_rows(condition, sep = "/") %>%
  separate_rows(condition, sep = " ")
```

##### Combine baseline data for HB, R2, R3
```{r baseline_combine}
bl <- bind_rows(bl_hb, bl_r2, bl_r3) %>%
  mutate_at(c("transect", "coral_id"), as.factor)
```

### Compliance monitoring (during construction)

Hardbottom, middle, and outer reef data are all in the same spreadsheet.

```{r load_data_nov2013_mar2015}
# Import data from file: NEW_Offshore_Comp_Scleractinian_Cond_All_Weeks_8_10_15-QAQC.xlsx
comp <- readxl::read_xlsx(
  path = "data/DEP_all/NEW_Offshore_Comp_Scleractinian_Cond_All_Weeks_8_10_15-QAQC.xlsx",
  sheet = "All Weeks", guess_max = 3000)

# Tidy data
comp <- comp %>%
  clean_names() %>%
  rename(condition = condition_code) %>%
  select(date, week, visit, site, transect, coral_id, species, condition) %>%
  mutate(date = as_date(date)) %>%
  mutate_at(c("transect", "coral_id"), as.factor)

# Correct typos and errors in date column (detected by looking at distinct date/week combinations)
comp <- comp %>%
  # 2014-11-25 should be 2013-11-25 because it is Week #1
  mutate(date = if_else(date == "2014-11-25", as_date("2013-11-25"), as_date(date))) %>%
  # 2013-04-02 should be 2014-04-02
  mutate(date = if_else(date == "2013-04-02", as_date("2014-04-02"), as_date(date))) %>%
  # Looks like a mistake from filling down Date in Excel for Surveyor MLR 
  # Change all Week 27 dates for R2SC1-RR visit 1 to May 23, 2014
  mutate(date = case_when(site == "R2SC1-RR" & week == 27 & visit == 1 ~ as_date("2014-05-23"),
                          TRUE ~ as_date(date))) %>%
  # 2020-05-29 should be 2014-05-29
  mutate(date = if_else(date == "2020-05-29", as_date("2014-05-29"), as_date(date))) %>%
  # Fix missing dates for HBS1-CP week 28
  mutate(date = case_when(site=="HBS1-CP" & week == 28 & is.na(date) ~ as_date("2014-05-29"),
                          TRUE ~ as_date(date)))

# Select columns
comp <- comp %>% select(date, site, transect, coral_id, species, condition)

# Check and make sure no missing dates
#comp %>% filter(is.na(date))

# Aggregate condition codes into single row for each observation
# comp <- comp %>%
#   group_by(date, site, transect, coral_id, species) %>%
#   summarise(condition = list(condition_code)) %>%
#   ungroup()
```

### Immediate post-construction

#### April - May 2015
```{r load_data_aprmay2015}
# Import data from file: NEW_Offshore_Compliance_Scleractinian_Cond_R2_R3_IMPACT_ASSESS_data TAGGED CORALS.xlsx
pc1 <- readxl::read_xlsx(
  path = "data/raw/2015data/NEW_Offshore_Compliance_Scleractinian_Cond_R2_R3_IMPACT_ASSESS_data TAGGED CORALS.xlsx",
  sheet = "IA TAGGED ONLY")

# Tidy data
pc1 <- pc1 %>%
  clean_names() %>%
  rename(condition = condition_code) %>%
  select(date, site, transect, coral_id, species, condition) %>%
  mutate(date = as_date(date)) %>%
  mutate_at(c("transect", "coral_id"), as.factor)

# Check and make sure no missing dates
#pc1 %>% filter(is.na(date))

# Aggregate condition codes into single row for each observation
# pc1 <- pc1 %>%
#   group_by(date, site, transect, coral_id, species) %>%
#   summarise(condition = list(condition_code)) %>%
#   ungroup()
```

#### June - July 2015
```{r load_data_junjuly2015}
# Import data from file: NEW_Offshore_Post_Construction_Data_Entry_ALL_Weeks_GOOD 151015.xlsx
pc2 <- readxl::read_xlsx(
  path = "data/raw/NEW_Offshore_Post_Construction_Data_Entry_ALL_Weeks_GOOD 151015.xlsx",
  sheet = "Raw Data", guess_max = 3000)

# Tidy data
pc2 <- pc2 %>%
  clean_names() %>%
  filter(tagged_coral_y_n == "Y") %>%           # Tagged corals only
  rename(condition = condition_code_qaqc) %>%
  select(date, site, transect, coral_id, species, condition) %>%
  mutate(date = as_date(date)) %>%
  mutate_at(c("transect", "coral_id"), as.factor)

# Check and make sure no missing dates
#pc2 %>% filter(is.na(date))

# Aggregate condition codes into single row for each observation
# pc2 <- pc2 %>%
#   group_by(date, site, transect, coral_id, species) %>%
#   summarise(condition = list(condition_code_qaqc)) %>%
#   ungroup()
```

### One year post-construction (Aug 2016)

```{r load_data_aug2016}
# Import data from file: permanent_sites_impact_assessment_corals_2016_CSI_QAQC.xlsx
pc3 <- readxl::read_xlsx(
  path = "data/raw/permanent_sites_impact_assessment_corals_2016_CSI_QAQC.XLSX",
  sheet = "Raw Data", guess_max = 3000, skip = 1)

# Tidy data
pc3 <- pc3 %>%
  clean_names() %>%
  filter(tagged_coral_y_n == "Y") %>%              # Tagged corals only
  rename(condition = condition_code_qaqc) %>%
  select(date, site, transect, coral_id, species, condition) %>%
  mutate(date = as_date(date)) %>%
  mutate_at(c("transect", "coral_id"), as.factor)

# Check and make sure no missing dates
#pc3 %>% filter(is.na(date))

# Aggregate condition codes into single row for each observation
# pc3 <- pc3 %>%
#   group_by(date, site, transect, coral_id, species) %>%
#   summarise(condition = list(condition_code_qaqc)) %>%
#   ungroup()
```

# Combine all tagged coral data
```{r merge_data}
# Combine all data
tagged <- bind_rows(bl, comp, pc1, pc2, pc3)
```

### Homogenize site, transect, coral ID, and species codes
```{r}
tagged <- tagged %>%
  mutate(site = str_extract(site, pattern = "[^-]+")) %>%           # Remove "-RR" etc.
  mutate(transect = str_extract(transect, pattern = ".$")) %>%      # Get transect as single digit
  unite(coral, site, transect, coral_id, sep = ".", remove = FALSE)

tagged <- tagged %>%
  mutate(species = case_when(
    species %in% c("MFOR") ~ "MDEC",
    species %in% c("MFER", "MALI", "Mycetophyllia ferox") ~ "MYCE",
    species %in% c("Orbicella annularis") ~ "OANN",
    species %in% c("stephanoceonia intersepta", "Stephanocoenia intersepta",
                   "Stephanocoenia intercepta", "stephanocoenia intersepta") ~ "SINT",
    species %in% c("Dichocoenia stokesii", "Dichocenia stokesii") ~ "DSTO",
    species %in% c("Solenastrea bournoni") ~ "SBOU",
    species %in% c("Montastraea cavernosa") ~ "MCAV",
    species %in% c("Siderastrea siderea") ~ "SSID",
    species %in% c("Meandrina meandrites") ~ "MMEA",
    species %in% c("Diploria clivosa") ~ "DCLI",
    species %in% c("Agaricia sp.") ~ "AGSP",
    species %in% c("Madracis formosa") ~ "MFOR",
    species %in% c("Madracis decactis") ~ "MDEC",
    species %in% c("Porites astreoides") ~ "PAST",
    species %in% c("Porites porites") ~ "PPOR",
    species %in% c("Siderastrea sp.") ~ "SIDSP",
    species %in% c("Oculina diffusa") ~ "ODIF",
    species %in% c("Diploria strigosa", "Diplora strigosa", "DSTR") ~ "PSTR",
    species %in% c("Diploria labyrinthiformis") ~ "DLAB",
    species %in% c("Colpophyllia natans", "Colophyllia natans") ~ "CNAT",
    species %in% c("Eusmilia fastigiata") ~ "EFAS",
    species %in% c("Orbicella faveolata") ~ "OFAV",
    species %in% c("Acropora cervicornis") ~ "ACER",
    species %in% c("Agaricia agaricites") ~ "AAGA",
    species %in% c("Agaricia lamarcki") ~ "ALAM",
    TRUE ~ as.character(species)))

# Remove row with missing data
tagged <- tagged %>% drop_na(date)
```

# Additional data QC

#### Check condition codes
```{r}
tagged %>%
  distinct(condition) %>%
  print(n = nrow(.))

tagged <- tagged %>%
  mutate(condition = recode(condition, `sed dust` = "SED"))
```


#### Check for conflicting species IDs
```{r}
spp_conflicts <- tagged %>% 
  distinct(coral, species) %>%
  group_by(coral) %>%
  filter(n() > 1) %>%
  arrange(coral) %>%
  summarise(conflicting_species = list(species))
```

183 of the tagged corals have conflicting species information.

#### Correct conflicting IDs by assigning most frequently noted species
```{r}
# Get most frequently used species ID for each coral
spp_ids <- tagged %>%
  group_by(coral) %>%
  count(species) %>%
  summarise(most_freq_species = species[which.max(n)])

# Replace species ID with most frequently used species ID for each coral
tagged <- tagged %>%
  left_join(spp_ids, by = "coral") %>%
  select(date, site, transect, coral_id, coral, most_freq_species, condition) %>%
  rename(species = most_freq_species) 
```

#### Change *A. cervicornis* WP records to WB

```{r}
tagged <- tagged %>%
  mutate(condition = if_else(species == "ACER", 
                             str_replace(condition, "WP", "WB"),
                             condition))
```

#### Add site metadata
```{r}
tagged <- tagged %>%
  mutate(channel = if_else(str_detect(site, "C"), "control", "channelside"),
            reef = str_sub(site, 1, 2),
             dir = str_sub(site, 3, 3)) %>%
  select(date, reef, dir, channel, site, transect, coral, species, condition)
```

#### Spread all conditions to boolean columns for each observation
```{r}
tagged_bool <- tagged %>% 
  mutate(hasCond = TRUE) %>%
  distinct() %>%
  spread(condition, hasCond, fill = FALSE)
```

#### Nest all condition codes for each observation
```{r}
tagged_nest <- tagged %>% nest(condition, .key = "condition")
```

#### Concatenate all condition codes for each obervation
```{r}
tagged_c <- tagged %>%
  group_by(date, reef, dir, channel, site, transect, coral, species) %>%
  summarise(condition = paste(condition, collapse = ", ")) %>%
  ungroup()
```

# Save tagged coral data as .RData
```{r}
save(tagged, tagged_bool, tagged_nest, tagged_c, file = "data/tagged.RData")
```

# Write concatenated data to file
```{r}
# Write all data to spreadsheet (STILL CONTAINS UNCORRECTED SPP CONFLICTS)
dd <- tagged_c %>% 
  arrange(coral, date) %>%
  select(-channel, -reef, -dir, -coral)

write.table(dd, file="data/tagged_c.txt", row.names = FALSE, quote = FALSE, sep = "\t")
```







```{r, include = FALSE, eval = FALSE}
#### Identify colonies that died after having WP
wpbeforedead <- alldata %>%
  group_by(coral) %>%
  summarise(WP = as.character(min(Date[hasCond("^WP$", Condition.Codes)])),
            DEAD = as.character(min(Date[hasCond("^DEAD$", Condition.Codes)]))) %>%
  drop_na() %>%
  mutate(WPbeforeDEAD = as.Date(WP) < as.Date(DEAD))

# For these corals that had WP before DEAD, change DEAD to WPDEAD
alldata[alldata$coral %in% wpbeforedead$coral & hasCond("DEAD", alldata$Condition.Codes), "Condition.Codes"] <- "WPDEAD"
```

